---
title: "Light Penertration Proxies"
author: "Chih-Lin Wei"
date: "`r Sys.Date()`"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dev="jpeg", dpi=300)
#library(devtools)
#devtools::install_github("MikkoVihtakari/ggOceanMapsData") # required by ggOceanMaps
```

```{r}
library(ggOceanMaps)
library(ggOceanMapsData)
library(ncdf4)
library(raster)
library(RColorBrewer)
library(dplyr)
library(plyr)
library(patchwork)
library(knitr)
library(kableExtra)
library(scales)
library(utils)

large <- theme(legend.title = element_text(size=15),
        legend.text = element_text(size=15),
        axis.title = element_text(size=15),
        axis.text = element_text(size=15),
        strip.text = element_text(size=15))

no_legend <- theme(legend.position = "none")
vert_x <- theme(axis.text.x = element_text(angle = 60))
letter_strip <- theme(strip.text = element_text(hjust = 0, vjust = 1.5, margin = margin(0.2,0,0.2,0, "cm"), size=15), strip.background = element_blank())

Spectral.colors <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
ex <- extent(c(-180, 180, 50, 90)) # Eye Index data extant

# Function to convert scientific notation to exponential notation
scientific_10 <- function(x) { parse(text=gsub("e", "%*%10^", scales::label_scientific()(x)))}

#  Stereographic map axis title
hx <- cbind(transform_coord(lon = seq(-135, 180, by = 45), lat = rep(48, 8), proj.out = 3995), label = c("135°W", "90°W", "45°W", "0°", "45°E", "90°E", "135°E", "180°"), hjust = c(rep(0.8, 3), 0, rep(0.2, 3), 0.5))
vx <-  cbind(transform_coord(lon = rep(180, 4), lat = c(60, 70, 80, 90), proj.out = 3995), label = c("60°N", "70°N", "80°N", "90°N"))

# Eye Index data
er <- read.csv("../csv/fulldataset-20240227.csv")
er <- cbind(transform_coord(lon=er$Longitude, lat=er$Latitude, proj.out = 3995), er)
# Show species eye ratio
er$ER <- cut(er$Species, breaks=c(0, 0.0001, 0.25, 0.5, 0.75, 1.1), right=FALSE, labels=c("0", "0-0.25", "0.25-0.5", "0.5-0.75", "0.75-1"))

# Download Arctic land shapfile
url <- "https://github.com/MikkoVihtakari/ggOceanMapsLargeData/raw/master/data/arctic_land.rda"
destfile <- paste(tempdir(), "arctic_land.rda", sep="//")
download.file(url, destfile)
```

Lee et al. (2005) developed a model for vertical transmittance of solar radiation in the upper ocean. The model uses total absorption,  particulate backscatter, water depth, and solar zenith angle as inputs. In this test, I used the total absorption and backscatter at 488 nm from the entire mission composite climatology to calculate the vertical transmittance (Tvis) for the visible spectrum at surface, 25 m, euphotic depth, and seafloor. Tvis is the ratio between downwelling solar radiation at the surface, Evis(0), and at depth, Evis(z). The total absorption, backscatter, and euphotic depth were downloaded from [OceanColor Web L3 browser](https://oceancolor.gsfc.nasa.gov/l3/order/). The seafloor bathymetry were downloaded from [ETOPO Global Relief Model](https://www.ncei.noaa.gov/products/etopo-global-relief-model). The complete references for Lee's algorithms can be found [here](https://oceancolor.gsfc.nasa.gov/products/special/#ZLEE).

# Data from OceanColor Web
## IOP Total Absorption at 488 nm (MODIS_Aqua Entire Mission Composite climatology)

```{r}
nc <- nc_open("../OceanColorData/MODIS_Aqua/IOP_a_488/AQUA_MODIS.20020704_20220930.L3m.CU.IOP.a_488.4km.nc")
ng <- ncvar_get(nc, varid = "a_488")
a488 <- raster(t(ng), xmn=-180, xmx=180, ymn=-90, ymx=90, crs = "+proj=longlat +datum=WGS84") %>% crop(ex)
nc_close(nc)

r <- projectRaster(a488, crs = projection(ggOceanMapsData::arctic_bathy))
df <- rasterToPoints(r) %>% as.data.frame

p <- basemap(limits=50, shapefiles = "Arctic")+ 
  geom_raster(data=df, aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(colours=Spectral.colors(100), trans="log10")

p_a488 <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_colour_viridis_d()+
  labs(fill=expression("a488"~(m^-1)), colour="Eye Index", title=expression("a"~~~"Total Absorption at 488 nm"))+
  guides(fill = guide_colourbar(order=1, barheight=10), colour = guide_legend(order=2))+
  coord_sf(clip = 'off')+
  theme(legend.key = element_rect(fill = "gray80", colour = NA))
```

## IOP Total Backscatter at 488 nm (MODIS_Aqua Entire Mission Composite climatology)

```{r}
nc <- nc_open("../OceanColorData/MODIS_Aqua/IOP_bb_488/AQUA_MODIS.20020704_20220930.L3m.CU.IOP.bb_488.4km.nc")
ng <- ncvar_get(nc, varid = "bb_488")
bb488 <- raster(t(ng), xmn=-180, xmx=180, ymn=-90, ymx=90, crs = "+proj=longlat +datum=WGS84") %>% crop(ex)
nc_close(nc)

r <- projectRaster(bb488, crs = projection(ggOceanMapsData::arctic_bathy))
df <- rasterToPoints(r) %>% as.data.frame

p <- basemap(limits=50, shapefiles = "Arctic")+ 
  geom_raster(data=df, aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(colours=Spectral.colors(100), trans="log10")

p_bb488 <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_color_viridis_d()+
  labs(fill=expression("bb488"~(m^-1)), colour="Eye Index", title=expression("b"~~~"Total Backscatter at 488 nm"))+
  guides(fill = guide_colourbar(order=1, barheight=10), colour = guide_legend(order=2))+
  coord_sf(clip = 'off')+
  theme(legend.key = element_rect(fill = "gray80", colour = NA))
```

```{r, eval=TRUE, fig.width=12, fig.height=5, fig.cap="Fig.S1. IOP total absorption at 488 nm and total backscatter at 488 nm based on MODIS_Aqua entire mission composite climatology."}
p_a488+p_bb488
```

## Euphotic depth (MODIS_Aqua Entire Mission Composite climatology) from Lee Algorithm (ZLEE)

```{r, fig.width=6, fig.height=5}
nc <- nc_open("../OceanColorData/MODIS_Aqua/Zeu_lee/A20021852022059.L3m_CU_ZLEE_Zeu_lee_4km.nc")
ng <- ncvar_get(nc, varid = "Zeu_lee")
Zeu <- raster(t(ng), xmn=-180, xmx=180, ymn=-90, ymx=90, crs = "+proj=longlat +datum=WGS84") %>% crop(ex)
nc_close(nc)

r <- projectRaster(Zeu, crs = projection(ggOceanMapsData::arctic_bathy))
df <- rasterToPoints(r) %>% as.data.frame

p <- basemap(limits=50, shapefiles = "Arctic")+ 
  geom_raster(data=df, aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(colours=Spectral.colors(100) %>% rev)

p_zeu <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_color_viridis_d()+
  labs(fill="Zeu (m)", colour="Eye Index", title=expression("a"~~~"Euphotic depth"))+
  guides(fill = guide_colourbar(order=1, barheight=10), colour = guide_legend(order=2))+
  coord_sf(clip = 'off')+
  theme(legend.key = element_rect(fill = "gray80", colour = NA))
```

## Photosynthetically available radiation (MODIS_Aqua Entire Mission Composite climatology)

```{r, fig.width=6, fig.height=5}
nc <- nc_open("../OceanColorData/MODIS_Aqua/PAR/AQUA_MODIS.20020704_20220930.L3m.CU.PAR.par.4km.nc")
ng <- ncvar_get(nc, varid = "par")
par <- raster(t(ng), xmn=-180, xmx=180, ymn=-90, ymx=90, crs = "+proj=longlat +datum=WGS84")  %>% crop(ex)
nc_close(nc)

r <- projectRaster(par, crs = projection(ggOceanMapsData::arctic_bathy))
df <- rasterToPoints(r) %>% as.data.frame
df$layer[df$layer>quantile(df$layer, .99)] <- quantile(df$layer, .99) # make 99 percentile as maximum value

p <- basemap(limits=50, shapefiles = "Arctic")+ 
  geom_raster(data=df, aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(colours=Spectral.colors(100))

p_par <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_color_viridis_d()+
  labs(fill=expression("PAR"~(E~m^-2*d^-1)), colour="Eye Index", title=expression("b"~~~"Photosynthetically available radiation"))+
  guides(fill = guide_colourbar(order=1, barheight=10), colour = guide_legend(order=2))+
  coord_sf(clip = 'off')+
  theme(legend.key = element_rect(fill = "gray80", colour = NA))
```

```{r, fig.width=12, fig.height=5, eval=FALSE}
p_zeu+p_par
```

## ETOPO 2022 Global Relief Model

```{r}
nc <- nc_open("../ETOPO2022/ETOPO_2022_v1_60s_N90W180_bed.nc")
ng <- ncvar_get(nc, varid = "z")
bathy <- raster(t(ng[,10800:1]), xmn=-180, xmx=180, ymn=-90, ymx=90, crs = "+proj=longlat +datum=WGS84")  %>% crop(ex)
nc_close(nc)

bathy <- calc(bathy, fun=function(x){x[x>0]<-NA; return(-x)})

bathy <- projectRaster(bathy, a488)

p <- basemap(limits=50, shapefiles = "Arctic", bathymetry=TRUE)

p_bathy <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_color_viridis_d()+
  labs(fill="Depth (m)", colour="Eye Index", title=expression("a"))+
  coord_sf(clip = 'off')

p <- basemap(limits=50, shapefiles = "Arctic", bathymetry=FALSE)

p_region <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=Region))+
  scale_color_viridis_d()+
  labs(fill="Depth (m)", colour="Region", title=expression("b"))+
  coord_sf(clip = 'off')
```

```{r, eval=TRUE, fig.width=12, fig.height=5, fig.cap="Figure 1. Distribution map showing the location and eye index (EIS) and region of each sample from the Arctic Ostracods Database in the Arctic Ocean."}
p_bathy+p_region
```

# Lee et al. (2005) [Journal of Geophysical Research, 110(C2)](http://dx.doi.org/10.1029/2004jc002275)
## Table 2. Values of the Model Parameters for Kvis(z)

```{r}
X0 = -0.057
X1 = 0.482
X2 = 4.221
sigma0 = 0.183
sigma1 = 0.702
sigma2 = -2.567
alpha0 = 0.090
alpha1 = 1.465
alpha2 = -0.667
```

## Vertical transmittance (Tvis) at surface, 25-m depth, and euphotic depth at zero solar zenith angle

```{r}
# Function for two-parameter model to estimate vertical transmittance (Tvis)
Tvis_fun <- function(a488, bb488, z, theta){
  # Equation 9a
  K1 = (X0 + X1 * a488^0.5 + X2 * bb488) * (1 + alpha0 * sin(theta * pi/180))
  # Equation 9b
  K2 = (sigma0 + sigma1 * a488 + sigma2 * bb488) * (alpha1 +alpha2 * cos(theta * pi/180))
  # Equation 7
  # Attenuation coefficient
  Kvis = K1 + K2/(1 + z)^0.5
  # Equation 10
  # Vertical transmittance
  Tvis <- exp(-Kvis * z)
  return(Tvis)
}
```

### Tvis at surface

```{r, fig.height=5, fig.width=8.5}
Tvis <- Tvis_fun(a488, bb488, 0, 0)

r <- projectRaster(Tvis, crs = projection(ggOceanMapsData::arctic_bathy))
df <- rasterToPoints(r) %>% as.data.frame

p <- basemap(limits=50, shapefiles = "Arctic")+ 
  geom_raster(data=df, aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(colours=Spectral.colors(100))

p_tvis0 <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_color_viridis_d()+
  labs(fill=expression(italic(T[vis](0))), colour="Eye Index", title=expression("a"~~~"Vertical transmittance at surface"))+
  guides(fill = guide_colourbar(order=1, barheight=10), colour = guide_legend(order=2))+
  coord_sf(clip = 'off')+
  theme(legend.key = element_rect(fill = "gray80", colour = NA))
```

### Tvis at 25 m

```{r, fig.height=5, fig.width=8.5}
Tvis <- Tvis_fun(a488, bb488, 25, 0)

r <- projectRaster(Tvis, crs = projection(ggOceanMapsData::arctic_bathy))
df <- rasterToPoints(r) %>% as.data.frame

p <- basemap(limits=50, shapefiles = "Arctic")+ 
  geom_raster(data=df, aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(colours=Spectral.colors(100))

p_tvis25 <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_color_viridis_d()+
  labs(fill=expression(italic(T[vis](25))), colour="Eye Index", title=expression("b"~~~"Vertical transmittance at 25-m"))+
  guides(fill = guide_colourbar(order=1, barheight=10), colour = guide_legend(order=2))+
  coord_sf(clip = 'off')+
  theme(legend.key = element_rect(fill = "gray80", colour = NA))
```

### Tvis at 25 m * PAR

```{r, fig.height=5, fig.width=8.5}
r <- projectRaster(Tvis*par, crs = projection(ggOceanMapsData::arctic_bathy))
df <- rasterToPoints(r) %>% as.data.frame

p <- basemap(limits=50, shapefiles = "Arctic")+ 
  geom_raster(data=df, aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(colours=Spectral.colors(100))

p_par25 <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_color_viridis_d()+
  labs(fill=expression("PAR"~(25)~("E"~m^-2*d^-1)), colour="Eye Index", title=expression("c"~~~"Photosynthetically available radiation at 25-m"))+
  guides(fill = guide_colourbar(order=1, barheight=10), colour = guide_legend(order=2))+
  coord_sf(clip = 'off')+
  theme(legend.key = element_rect(fill = "gray80", colour = NA))
```

### Tvis at euphotic depth

```{r, fig.height=5, fig.width=8.5}
Tvis <- Tvis_fun(a488, bb488, Zeu, 0)
# remove the top 1% of data
Tvis[Tvis>quantile(Tvis, 0.99)] <- NA

r <- projectRaster(Tvis, crs = projection(ggOceanMapsData::arctic_bathy))
df <- rasterToPoints(r) %>% as.data.frame

p <- basemap(limits=50, shapefiles = "Arctic")+ 
  geom_raster(data=df, aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(colours=Spectral.colors(100))

p_tvis_zeu <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_color_viridis_d()+
  labs(fill=expression(italic(T[vis](Z[eu]))), colour="Eye Index", title=expression("d"~~~"Vertical transmittance at Zeu"))+
  guides(fill = guide_colourbar(order=1, barheight=10), colour = guide_legend(order=2))+
  coord_sf(clip = 'off')+
  theme(legend.key = element_rect(fill = "gray80", colour = NA))
```

```{r, fig.width=12, fig.height=10, eval=FALSE}
(p_tvis0+p_tvis25)/(p_par25+p_tvis_zeu)
```

```{r, fig.width=12, fig.height=5, fig.cap="Fig. 5  Geographical distribution of light availability proxies. Heat maps indicate (A) Tvis (Vertical transmittance) at 25-m water depth and (B) PAR (Photosynthetically available radiation). Note that sites in the high Arctic with white background indicates permanently ice-covered sites and thus no light availability data is available for those sites. Eye index (EIS) shown on the maps."}
p <- p_tvis25+labs(title=expression("a"~~~"Vertical transmittance at 25-m"))
p+p_par
```

### Tvis at bottom

```{r, fig.height=5, fig.width=8.5}
Tvis <- Tvis_fun(a488, bb488, bathy, 0)

r <- projectRaster(Tvis, crs = projection(ggOceanMapsData::arctic_bathy))
df <- rasterToPoints(r) %>% as.data.frame

p <- basemap(limits=50, shapefiles = "Arctic")+ 
  geom_raster(data=df, aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(colours=Spectral.colors(100), trans= trans_new("5rt", function(x)x^(0.2), function(x)x^5))

p_tvis_bottom <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_color_brewer(type="seq", palette="YlOrBr", direction=-1)+
  labs(fill=expression(italic(T[vis](z))), colour="Eye Index", title=expression("a"~~~"Vertical transmittance at bottom"))+
  guides(fill = guide_colourbar(order=1, barheight=10), colour = guide_legend(order=2))+
  coord_sf(clip = 'off')+
  theme(legend.key = element_rect(fill = "gray80", colour = NA))
```

### Tvis at bottom * PAR

```{r, fig.height=5, fig.width=8.5}
r <- projectRaster(Tvis*par, crs = projection(ggOceanMapsData::arctic_bathy))
df <- rasterToPoints(r) %>% as.data.frame

p <- basemap(limits=50, shapefiles = "Arctic")+ 
  geom_raster(data=df, aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(colours=Spectral.colors(100), trans= trans_new("5rt", function(x)x^(0.2), function(x)x^5))

p_par_bottom <- reorder_layers(p)+
  geom_text(data=hx, aes(x=lon, y=lat, label=label, hjust=hjust), colour="gray30")+
  geom_text(data=vx, aes(x=lon, y=lat, label=label), colour="gray30")+
  geom_point(data=er, aes(x=lon, y=lat, colour=ER))+
  scale_color_brewer(type="seq", palette="YlOrBr", direction=-1)+
  labs(fill=expression("PAR"~(italic(z))~("E"~m^-2*d^-1)), colour="Eye Index", title=expression("b"~~~"Photosynthetically available radiation at bottom"))+
  guides(fill = guide_colourbar(order=1, barheight=10), colour = guide_legend(order=2))+
  coord_sf(clip = 'off')+
  theme(legend.key = element_rect(fill = "gray80", colour = NA))
```

```{r, fig.width=12, fig.height=5, fig.cap="Figure 5 Geographical distribution of light availability proxies. Heat maps indicate (A) Vertical transmittance and (B) Photosynthetically available radiation at sampling depth [Tvis (z) and PAR (z), respectively]. Note that for sites in the high Arctic the white background indicates permanently ice-covered sites and thus no light availability data is available for those sites. Eye index shown on the maps."}
p_tvis_bottom+p_par_bottom
```

## Extract a488, bb488, Zeu and calculate TVis

We extracted the gridded a488 and bb488 by sampling coordinates. For the site outside of the grid coverage, we extract the mean grid values within a 100-km radius buffer. The remaining sites not matched by spatial grids or buffers are under the permanent sea ice cover (between 84.3 and 89.98 latitudes); therefore, their Tvis are set to zero.

```{r}
# Data distribution
loc <- er[, c("Longitude", "Latitude")]
coordinates(loc) <- c("Longitude", "Latitude")
projection(loc) <- "+proj=longlat +datum=WGS84"

# IOP Total Absorption at 488 nm (m^-1)
ta <- extract(a488, loc)
na <- is.na(ta)
mi <- extract(a488, loc[na,], buffer=100000, fun=mean, na.rm=TRUE)
ta0 <- ta
ta0[na] <- mi

# IOP Total Backscatter at 488 nm (m^-1)
tb <- extract(bb488, loc)
na <- is.na(tb)
mi <- extract(bb488, loc[na,], buffer=100000, fun=mean, na.rm=TRUE)
tb0 <- tb
tb0[na] <- mi

# Euphotic depth (m)
ed <- extract(Zeu, loc)
na <- is.na(ed)
mi <- extract(Zeu, loc[na,], buffer=100000, fun=mean, na.rm=TRUE)
ed0 <- ed
ed0[na] <- mi

# Photosynthetically available radiation (einstein m^-2 day^-1)
pa <- extract(par, loc)
na <- is.na(pa)
mi <- extract(par, loc[na,], buffer=100000, fun=mean, na.rm=TRUE)
pa0 <- pa
pa0[na] <- mi

#TVIS at sampling depth
tv <- Tvis_fun(ta, tb, er$Water.Depth, 0)
tv0 <- Tvis_fun(ta0, tb0, er$Water.Depth, 0)
# All the missing values are under permanent ice cover
#coordinates(loc[is.na(tv0),])[,2] %>% as.vector %>% range
tv0[is.na(tv0)] <- 0

out <- cbind(er, "a488"= ta, "bb488" = tb, "Zeu" = ed, "PAR" = pa, "Tvis" = tv, "a488_buffer"= ta0, "bb488_buffer" = tb0, "Zeu_buffer" = ed0, "PAR_buffer" = pa0, "Tvis_buffer" = tv0)
out$Tvis_PAR <- with(out, Tvis_buffer*PAR_buffer)

with(out, cbind(Water.Depth, Latitude))%>%cor(use="pairwise.complete.obs")

with(subset(out, Tvis_PAR>0), cbind(Water.Depth, Latitude, log_Tvis_PAR=log(Tvis_PAR), log_Tvis=log(Tvis_buffer)))%>%cor(use="pairwise.complete.obs")

with(subset(out, Tvis_PAR>0), cbind(Tvis_buffer, Tvis_PAR))%>%cor(use="pairwise.complete.obs")

write.csv(out, file= "../csv/fulldataset_Tvis.csv", row.names=FALSE)
```

```{r}
library(ggplot2)
library(reshape2)
library(car)
library(mgcv)
library(doBy)
library(lme4)

ml <- melt(out, id.vars=c("Number", "Water.Depth", "Latitude", "Longitude", "Region", "Tvis", "Zeu", "PAR", "Total"), measure.vars = c("Abundance", "Species", "Genus"), value.name = "Eye_Ratio")
ml$Tvis_PAR <- with(ml, Tvis*PAR)
ml$variable <- factor(ml$variable, labels=c("Individual", "Species", "Genus"))

ml0 <- melt(out, id.vars=c("Number", "Water.Depth", "Latitude", "Longitude", "Region", "Tvis_buffer", "Zeu_buffer", "PAR_buffer", "Total"), measure.vars = c("Abundance", "Species", "Genus"), value.name = "Eye_Ratio")
ml0$Tvis_PAR <- with(ml0, Tvis_buffer*PAR_buffer)
ml0$variable <- factor(ml0$variable, labels=c("Individual", "Species", "Genus"))
ml0$withEye <- with(ml0, round(Total*Eye_Ratio, 0))
ml0$noEye <- with(ml0, Total-withEye)
```

# Fitting water depth to eye indices
## Generalized Linear Mixed-Effects Models (GLMM)

```{r}
# Eye indices as response variables
# Water depth as fixed factor (scaled due to large depth differences)
# Region as random factor

s1 <- lapply(splitBy(~variable, ml0)[2], FUN=function(x)glmer(Eye_Ratio~Water.Depth + (1 | Region), family = binomial, data=x) %>% summary)

s2 <- lapply(splitBy(~variable, ml0)[2], FUN=function(x)glmer(Eye_Ratio~Tvis_buffer+(1|Region), family = binomial, data=x) %>% summary)

s3 <- lapply(splitBy(~variable, ml)[2], FUN=function(x)glmer(Eye_Ratio~Tvis_PAR+(1|Region), family = binomial, data=x) %>% summary)

s4 <- lapply(splitBy(~variable, ml0)[2], FUN=function(x)glmer(Eye_Ratio~Latitude+(1|Region), family = binomial, data=x) %>% summary)

rbind(s1$Species$coefficients, NA, s2$Species$coefficients, NA,s3$Species$coefficients, NA,s4$Species$coefficients)%>%kable(caption="Table x. GLMM with binomial distribution on eye indices with water depth, Tvis(z), PAR(z), or latitude as a fixed factor and region as a random factor") %>% kable_classic(full_width = F, html_font = "Cambria", font_size=16)
```

```{r}
# According to Crawley (2012) The R Book Chapter 16: Propositional data. https://onlinelibrary.wiley.com/doi/10.1002/9781118448908.ch16
# Combine numbers of specimens with eyes and without eye as two-vector response variable.
# Water depth as fixed factor (scaled due to large depth differences)
# Region as random factor

s1 <- lapply(splitBy(~variable, ml0)[2], FUN=function(x){
  y <- cbind(x$withEye, x$noEye)
  glmer(y~Water.Depth + (1 | Region), family = binomial, data=x) %>% summary
  })

s2 <- lapply(splitBy(~variable, ml0)[2], FUN=function(x){
  y <- cbind(x$withEye, x$noEye)
  glmer(y~Tvis_buffer + (1 | Region), family = binomial, data=x) %>% summary
  })

s3 <- lapply(splitBy(~variable, ml0)[2], FUN=function(x){
  y <- cbind(x$withEye, x$noEye)
  glmer(y~Tvis_PAR + (1 | Region), family = binomial, data=x) %>% summary
  })

s4 <- lapply(splitBy(~variable, ml0)[2], FUN=function(x){
  y <- cbind(x$withEye, x$noEye)
  glmer(y~Latitude + (1 | Region), family = binomial, data=x) %>% summary
  })

rbind(s1$Species$coefficients, NA, s2$Species$coefficients, NA,s3$Species$coefficients, NA,s4$Species$coefficients)%>%kable(caption="Table x. GLMM with binomial distribution on the numbers of specimen with eyes and without eye using water depth, Tvis(z), PAR(z), or latitude as a fixed factor and region as a random factor") %>% kable_classic(full_width = F, html_font = "Cambria", font_size=16)
```

```{r}
ml0$Lab <- factor(ml0$variable, labels = c(NA, "a", NA))

# Menzies, George & Rowe (1968) Nature 217:93-95
# Isopod eye ratio @ 65 degree N
Water.Depth = c(2, 20, 200, 1000, 2000, 4000, 6000)
Eye_Ratio <- c(1, .65, .46, .17, .08, .02, 0)
iso <- data.frame(Water.Depth, Eye_Ratio, variable = "Species", Lab = "a")

p_depth_eye1 <- ggplot(data=subset(ml0, variable=="Species"), aes(x=Water.Depth, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(data=subset(ml0, variable=="Species"), method="glm", formula = y~x, method.args = list(family = "binomial"))+
  geom_point(data=iso, aes(x=Water.Depth, y=Eye_Ratio), colour = "red")+
  geom_path(data=iso, aes(x=Water.Depth, y=Eye_Ratio), colour = "red", linetype=2)+
  coord_flip()+
  scale_x_reverse()+
  scale_colour_viridis_d()+
  labs(x="Water depth (m)", y="Eye Index")+
  facet_wrap(~Lab, scales="free", nrow=2)+
  theme_bw() %+replace% large %+replace% vert_x %+replace% letter_strip 
```

```{r}
ml0$Lab <- factor(ml0$variable, labels = c(NA, "b", NA))
iso$Lab <- "b"

p_depth_eye2 <- ggplot(data=subset(ml0, variable=="Species"), aes(x=Water.Depth, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(method="glm", formula = y~x, method.args = list(family = "binomial"))+
  geom_point(data=iso, aes(x=Water.Depth, y=Eye_Ratio), colour = "red")+
  geom_path(data=iso, aes(x=Water.Depth, y=Eye_Ratio), colour = "red", linetype=2)+
  coord_flip(xlim = c(500, 0))+
  scale_x_reverse()+
  scale_colour_viridis_d()+
  labs(x="Water depth (m)", y="Eye Index")+
  facet_wrap(~Lab, scales="free", nrow=2)+
  theme_bw() %+replace% large %+replace% vert_x %+replace% letter_strip 
```

```{r, fig.width=10, fig.height=5, fig.cap="Figure 3. Bathymetric distribution of eye index values in the Arctic Ocean. Panel a shows the entire depth range, and panel b only shows the samples from the shallow region (0-500 m). The blue line shows significant relationship with 95% confidence interval (shaded area) based on the Generalized Linear Mixed-Effect Model (GLMM) with binomial distribution. Red symbols and dashed lines show isopod eye index from Menzies et al. (1968)."}
p1 <- p_depth_eye1+labs(y="Eye Index") + no_legend
p2 <- p_depth_eye2+theme(axis.title.y = element_blank()) 
p1+p2
```

```{r, fig.width=10, fig.height=8, fig.cap="Figure 4. Bathymetric distribution of eye index values computed by species method (EIS) in different regions (A–F) of the Arctic Ocean. Only samples from relatively shallow regions (0-500 m) are presented to show the regional difference of eye index in shallow water among the various Arctic Ocean coastal seas."}
eii_r <- subset(ml0, variable =="Species")
eii_r$Lab <- eii_r$Region
eii_r$Lab <- factor(eii_r$Lab, levels = c("Chukchi", "Beaufort", "Baffin", "Greenland", "Barents", "Kara", "Laptev", "Siberian"), labels = c("a   Chukchi", "b   Beaufort", "c   Baffin", "d   Greenland", "e   Barents", "f   Kara", "g   Laptev", "h   Siberian"))
ggplot(data=eii_r, aes(x=Water.Depth, y=Eye_Ratio))+
  geom_point()+
  coord_flip(xlim = c(500, 0))+
  scale_x_reverse()+
  scale_colour_viridis_d()+
  labs(x="Water depth (m)", y="Eye Index")+
  facet_wrap(~Lab, nrow=2)+
  theme_bw() %+replace% large %+replace% vert_x %+replace% letter_strip 
```


## Scatter plots with Tvis at Sampling Depth

```{r, eval=FALSE}
s <- lapply(splitBy(~variable, ml)[2], FUN=function(x)glmer(Eye_Ratio~Tvis+(1|Region), family = binomial, data=x) %>% summary)

lapply(s, FUN=function(x)cbind(Variable=rownames(x$coef), data.frame(x$coef)))%>%ldply%>%kable(caption="Table x. GLMM with binomial distribution on eye indices with water depth as fixed factor and regions as random factor") %>% kable_classic(full_width = F, html_font = "Cambria", font_size=16)

ml$Lab <- factor(ml$variable, labels = c("A   Individual", "B   Species", NA))

ggplot(data=subset(ml, variable=="Species"), aes(x=Tvis, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(data=subset(ml, variable=="Species"), method="glm", formula = y~x, method.args = list(family = "binomial"))+
  scale_colour_viridis_d()+
  labs(x=expression("Vertical transmittance"~italic(T[vis](z))~"at sampling depth"), y="Eye Index")+
  #facet_wrap(~Lab, scales="free")+
  theme_bw() %+replace% large %+replace% vert_x %+replace% letter_strip
```

```{r}
ml0$Lab <- factor(ml0$variable, labels = c(NA, "a", NA))

p_Tvis_eye <- ggplot(data=subset(ml0, variable=="Species"), aes(x=Tvis_buffer, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(data=subset(ml0, variable=="Species"), method="glm", formula = y~x, method.args = list(family = "binomial"))+
  scale_colour_viridis_d()+
  coord_flip()+
  scale_x_log10(labels = scientific_10)+
  labs(x=expression("Vertical transmittance"~italic(T[vis](z))~"at sampling depth with 100-km Buffer"), y="Eye Index")+
  facet_wrap(~Lab, scales="free")+
  theme_bw() %+replace% large %+replace% vert_x %+replace% letter_strip
```

## Scatter plots with PAR at Sampling Depth
```{r, eval=FALSE}
ml$Lab <- factor(ml$variable, labels = c("A   Individual", "B   Species", NA))

# E m-2 d-1 = 6.97 x 10^12 μmol quanta m-2 s-1
ggplot(data=subset(ml, variable=="Species"), aes(x=Tvis*PAR*6.97E12, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(data=subset(ml, variable!="Genus"), method="glm", formula = y~x, family = binomial, method.args = list(family = "binomial"))+
  coord_flip()+
  scale_x_log10(labels = scientific_10)+
  scale_colour_viridis_d()+
  labs(x=expression("PAR"~(italic(z))~(mu*mol~"quanta"~m^-2*s^-1)), y="Eye Index")+
  #facet_wrap(~Lab, scales="free")+
  theme_bw() %+replace% large %+replace% letter_strip
```

```{r}
ml0$Lab <- factor(ml0$variable, labels = c(NA, "b", NA))

# E m-2 d-1 = 6.97 x 10^12 μmol quanta m-2 s-1
p_PAR_eye <- ggplot(data=subset(ml0, variable=="Species"), aes(x=Tvis_PAR*6.97E12, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(data=subset(ml0, variable=="Species"), method="glm", formula = y~x, family = binomial, method.args = list(family = "binomial"))+
  coord_flip()+
  scale_x_log10(labels = scientific_10)+
  scale_colour_viridis_d()+
  labs(x="Photosynthetically available radiation (μmol quanta m^-2 s^-1) at sampling depth with 100-km Buffer", y="Eye Index")+
  facet_wrap(~Lab, scales="free")+
  theme_bw() %+replace% large %+replace% letter_strip
```

```{r, fig.width=10, fig.height=8, fig.cap="Figure 6 Eye index and light availability proxies of the vertical transmittance and photosynthetically available radiation at sampling depth [Tvis (z) and PAR (z), respectively]. The blue line shows significant relationship with 95% confidence interval (shaded area) based on Generalized Linear Mixed-Effect Model (GLMM) with binomial distribution."}
p1 <- p_Tvis_eye+labs(x=expression("Vertical transmittance"~italic(T[vis](z))), y="Eye Index")
p2 <- p_PAR_eye+labs(x=expression("PAR"~(italic(z))~(mu*mol~"quanta"~m^-2*s^-1)), y="Eye Index") + no_legend
p1/p2
```

```{r, fig.cap="Figure 7 Latitudinal gradient of eye index. The blue line shows significant relationship with 95% confidence interval (shaded area) based on Generalized Linear Mixed-Effect Model (GLMM) with binomial distribution."}
ml0$Lab <- factor(ml0$variable, labels = c("A   Individual", "B   Species", NA))

ggplot(data=subset(ml0, variable=="Species"), aes(x=Latitude, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(method="glm", formula = y~x, method.args = list(family = "binomial"))+
  scale_colour_viridis_d()+
  labs(x=expression("Latitude"~(degree*N)), y="Eye Index")+
  #facet_wrap(~Lab, scales="free")+
  theme_bw() %+replace% large %+replace% letter_strip
```

```{r}
# Eye indices as response variables
# Water depth and latitude as fixed factors 
# Region as random factor

s <- lapply(splitBy(~variable, ml0)[2], FUN=function(x)glmer(Eye_Ratio~Water.Depth + Latitude + (1 | Region), family = binomial, data=x) %>% summary)

lapply(s, FUN=function(x)cbind(Variable=rownames(x$coef), data.frame(x$coef)))%>%ldply%>%kable(caption="Table x. GLMM with binomial distribution on eye indices with water depth and latitude as fixed factors and regions as random factor") %>% kable_classic(full_width = F, html_font = "Cambria", font_size=16)
```

```{r, fig.width=10, fig.height=8, fig.cap="Fig. x. Eye index (EIS) as functions of the sampling depth and latitude. EIS is computed by species method."}

ml0$Lab <- factor(ml0$variable, labels = c(NA, "a", NA))

p1 <- ggplot(data=subset(ml0, variable=="Species"), aes(x=Water.Depth, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(data=subset(ml0, variable=="Species"), method="glm", formula = y~x, method.args = list(family = "binomial"))+
  coord_flip()+
  scale_x_reverse()+
  scale_colour_viridis_d()+
  labs(x="Water depth (m)", y="Eye Index")+
  facet_wrap(~Lab, scales="free")+
  theme_bw() %+replace% large %+replace% letter_strip

ml0$Lab <- factor(ml0$variable, labels = c(NA, "b", NA))

p2 <- ggplot(data=subset(ml0, variable=="Species"), aes(x=Latitude, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(data=subset(ml0, variable=="Species"), method="glm", formula = y~x, method.args = list(family = "binomial"))+
  coord_flip()+
  scale_colour_viridis_d()+
  labs(x=expression("Latitude"~(degree*N)), y="Eye Index")+
  facet_wrap(~Lab, scales="free")+
  theme_bw() %+replace% large %+replace% letter_strip %+replace% no_legend

p1/p2
```

```{r}
# Eye indices as response variables
# Latitude (scaled due to large depth differences) and PAR(z) as fixed factors 
# Region as random factor

s <- lapply(splitBy(~variable, subset(ml0, Tvis_PAR>0))[2], FUN=function(x)glmer(Eye_Ratio~Tvis_PAR + Latitude + (1 | Region), family = binomial, data=x) %>% summary)

lapply(s, FUN=function(x)cbind(Variable=rownames(x$coef), data.frame(x$coef)))%>%ldply%>%kable(caption="Table x. GLMM with binomial distribution on eye indices with photosynthetically available radiation (PAR) at sampling depth and latitude as fixed factors and regions as random factor") %>% kable_classic(full_width = F, html_font = "Cambria", font_size=16)
```

```{r, fig.width=10, fig.height=8, fig.cap="Fig. x. Eye index (EIS) as functions of the photosynthetically available radiation (PAR) at sampling depth and latitude. Panel (A), (C) are computed by abundance method; Panel (B), (D) are computed by species method."}

ml0$Lab <- factor(ml0$variable, labels = c(NA, "a", NA))

# E m-2 d-1 = 6.97 x 10^12 μmol quanta m-2 s-1
p1 <- ggplot(data=subset(ml0, variable=="Species"), aes(x=Tvis_PAR*6.97E12, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(data=subset(ml0, variable=="Species"), method="glm", formula = y~x, method.args = list(family = "binomial"))+
  coord_flip()+
  scale_x_log10(labels = scientific_10)+
  scale_colour_viridis_d()+
  labs(x=expression("PAR"~(italic(z))~(mu*mol~"quanta"~m^-2*s^-1)), y="Eye Index")+
  facet_wrap(~Lab, scales="free")+
  theme_bw() %+replace% large %+replace% letter_strip

ml0$Lab <- factor(ml0$variable, labels = c(NA, "b", NA))

p2 <- ggplot(data=subset(ml0, variable=="Species"), aes(x=Latitude, y=Eye_Ratio))+
  geom_point(aes(colour=Region))+
  stat_smooth(data=subset(ml0, variable=="Species"), method="glm", formula = y~x, method.args = list(family = "binomial"))+
  coord_flip()+
  scale_colour_viridis_d()+
  labs(x=expression("Latitude"~(degree*N)), y="Eye Index")+
  facet_wrap(~Lab, scales="free")+
  theme_bw() %+replace% large %+replace% letter_strip %+replace% no_legend

p1/p2
```

```{r}
summary(glmer(Tvis_buffer~Water.Depth+(1|Region), family = gaussian(link = "log"), data=subset(out, Tvis_buffer>0.001)))$coef %>% kable(caption = "Table x. GLMM with gaussian distribution and logarithmic link function on Tvis. Data with Tvis < 0.1% were deleted before analysis.") %>% kable_classic(full_width = F, html_font = "Cambria", font_size=16)

p_Tvis_depth <- ggplot(data=subset(out, Tvis_buffer!=0), aes(x=Water.Depth, y=Tvis_buffer))+
  geom_point(aes(colour=Region))+
  stat_smooth(method="glm", formula = y~x, method.args = list(family = gaussian(link = "log")))+
  coord_flip()+
  scale_x_reverse()+
  scale_y_continuous(labels = scientific_10)+
  scale_colour_viridis_d()+
  labs(x=expression("Water depth"~(m)), y=expression(italic(T[vis](z))~"at sampling depth"), tag = "a")+
  theme_bw() %+replace% large %+replace% vert_x 
```

```{r, fig.width=10, fig.height=4}
summary(glmer(Tvis_PAR~Water.Depth+(1|Region), family = gaussian(link = "log"), data=subset(out, Tvis_buffer>0.001)))$coef%>% kable(caption = "Table x. GLMM with gaussian distribution and logarithmic link function on PAR(z) at sampling depth. Data with Tvis < 0.1% were deleted before analysis.") %>% kable_classic(full_width = F, html_font = "Cambria", font_size=16)

# E m-2 d-1 = 6.97 x 10^12 μmol quanta m-2 s-1
p_PAR_depth <- ggplot(data=subset(out, Tvis_buffer!=0), aes(x=Water.Depth, y=Tvis_PAR*6.97E12))+
  geom_point(aes(colour=Region))+
  stat_smooth(method="glm", formula = y~x, method.args = list(family = gaussian(link = "log")))+
  coord_flip()+
  scale_x_reverse()+
  scale_y_continuous(labels = scientific_10)+
  scale_colour_viridis_d()+
  labs(x=expression("Water depth"~(m)), y=expression("PAR"~(italic(z))~(mu*mol~"quanta"~m^-2*s^-1)), tag="b")+
  theme_bw() %+replace% large %+replace% vert_x 
```

```{r, fig.width=12, fig.height=6, fig.cap="Fig. X. Vertical transmittance (Tvis) and photosynthetically available radiation (PAR) at sampling depth as functions of water depth. Data with Tvis = 0% were removed before conducting GLMM with with gaussian distribution and logarithmic link"}
p1 <- p_Tvis_depth + labs(y=expression(italic(T[vis](z)))) + no_legend
p2 <- p_PAR_depth + labs(y=expression("PAR"~(italic(z))~(mu*mol~"quanta"~m^-2*s^-1))) 
p1+p2
```

```{r, fig.width=12, fig.height=6, fig.cap="Fig. 8. Vertical transmittance (Tvis) and photosynthetically available radiation (PAR) at sampling depth as functions of water depth. Data with Tvis = 0% were removed before conducting GLMM with with gaussian distribution and logarithmic link"}
p1_log <- p1+scale_y_log10(labels = scientific_10) + stat_smooth(method="glm", formula = y~x)
p2_log <- p2+scale_y_log10(labels = scientific_10) + stat_smooth(method="glm", formula = y~x)
p1_log+p2_log
```

```{r}
summary(glmer(Tvis_buffer~Water.Depth+(1|Region), family = gaussian(link = "log"), data=subset(out, Tvis_buffer>0.001)))$coef %>% kable(caption = "Table x. GLMM with gaussian distribution and logarithmic link function on Tvis. Data with Tvis < 0.1% were deleted before analysis.") %>% kable_classic(full_width = F, html_font = "Cambria", font_size=16)

p_Tvis_depth <- ggplot(data=subset(out, Tvis_buffer>0.001), aes(x=Water.Depth, y=Tvis_buffer))+
  geom_point(aes(colour=Region))+
  stat_smooth(method="glm", formula = y~x, method.args = list(family = gaussian(link = "log")))+
  coord_flip()+
  scale_x_reverse()+
  scale_colour_viridis_d()+
  labs(x=expression("Water depth"~(m)), y=expression(italic(T[vis](z))~"at sampling depth"), tag = "a")+
  theme_bw() %+replace% large %+replace% vert_x 
```

```{r, fig.width=10, fig.height=4}
summary(glmer(Tvis_PAR~Water.Depth+(1|Region), family = gaussian(link = "log"), data=subset(out, Tvis_buffer>0.001)))$coef%>% kable(caption = "Table x. GLMM with gaussian distribution and logarithmic link function on PAR(z) at sampling depth. Data with Tvis < 0.1% were deleted before analysis.") %>% kable_classic(full_width = F, html_font = "Cambria", font_size=16)

# E m-2 d-1 = 6.97 x 10^12 μmol quanta m-2 s-1
p_PAR_depth <- ggplot(data=subset(out, Tvis_buffer>0.001), aes(x=Water.Depth, y=Tvis_PAR*6.97E12))+
  geom_point(aes(colour=Region))+
  stat_smooth(method="glm", formula = y~x, method.args = list(family = gaussian(link = "log")))+
  coord_flip()+
  scale_x_reverse()+
  scale_y_continuous(labels = scientific_10)+
  scale_colour_viridis_d()+
  labs(x=expression("Water depth"~(m)), y=expression("PAR"~(italic(z))~(mu*mol~"quanta"~m^-2*s^-1)), tag="a")+
  theme_bw() %+replace% large %+replace% vert_x 
```

```{r, fig.width=10, fig.height=10, fig.cap="Fig. 8 Photosynthetically available radiation at sampling depth [PAR (z)] as functions of water depth. (A) for the top 70 m; (B) for full bathymetric range. Data with Tvis (z) less than 0.1% were removed before conducting GLMM with gaussian distribution and logarithmic link function."}

p2 <- p_PAR_depth + labs(y=expression("PAR"~(italic(z))~(mu*mol~"quanta"~m^-2*s^-1))) 

p4 <- p2_log + labs(tag="b") + no_legend

p2/p4
```

